Computing Exact Restricted Isometry Constants via Mixed Integer Semidefinite Programming

### 2016-03-20 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: changed notation in Pataki's result

\noindent
In this section we propose an MISDP formulation to compute the optimal
constants $\alpha_k$ and $\beta_k$ in \eqref{lowerRIP} and
\eqref{upperRIP}, respectively. We will again start with the problem~\eqref{QP}.
For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$, we
again use semidefinite lifting, introducing the new variable $X=xx\T$ like in the last section. 
The cardinality-constraint in \eqref{QP} can equivalently be written using
binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i
\leq k$. We can thus equivalently rewrite \eqref{QP} as
\begin{equation}\label{Rk1MISDP}
 \begin{aligned}
  \min \quad & \Tr(A\T A X) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \\
  & \sum_{i=1}^n z_i \leq k,\\
  & \Rk(X) = 1, \\
  & X \succeq 0, \\
  & z \in \{0,1\}^n.
 \end{aligned}
\end{equation}
We then relax the non-convex rank constraint to arrive at the following
MISDP formulation:
\begin{equation}\label{MISDP}\tag{MISDP}
 \begin{aligned}
  \min \quad & \Tr(A\T A X) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \\
  & \sum_{i=1}^n z_i \leq k, \\
  & X \succeq 0, \\
  & z \in \{0,1\}^n.
 \end{aligned}
\end{equation}
In the following, we will show that \eqref{MISDP} actually produces the
exact value of $\alpha_k$. We will use the following result about the
relationship between the rank of solutions of semidefinite programs and the
dimension of corresponding faces of the feasible set~$P$, i.e., subsets $F
\subseteq P$ such that $x \in F$, $y, z \in P$, and $x = \tfrac{1}{2} (y + z)$ imply
$y,z \in F$:

\begin{theorem}[Pataki~\cite{pat98}]\label{patakiLemma}
  Let $X \in F$, where $F$ is a face of
  \begin{equation*}
    P \define \{X \in S_n \suchthat X \succeq 0,\; \Tr(A_i X) = b_i\; \text{ for } i = 1, \dots, \tilde{m}\}
  \end{equation*}
  for symmetric matrices $A_i \in S_n$, $i = 1, \dots, \tilde{m}$. Then
  \begin{equation*}
    \tfrac{1}{2}\Rk(X)\cdot(\Rk(X)+1) \leq \tilde{m} + dim(F).
  \end{equation*}
\end{theorem}

>>>>

\noindent
In this section we propose an MISDP formulation to compute the optimal
constants $\alpha_k$ and $\beta_k$ in \eqref{lowerRIP} and
\eqref{upperRIP}, respectively. We will again start with the problem~\eqref{QP}.
For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$, we
again use semidefinite lifting, introducing the new variable $X=xx\T$ like in the last section. 
The cardinality-constraint in \eqref{QP} can equivalently be written using
binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i
\leq k$. We can thus equivalently rewrite \eqref{QP} as
\begin{equation}\label{Rk1MISDP}
 \begin{aligned}
  \min \quad & \Tr(A\T A X) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \\
  & \sum_{i=1}^n z_i \leq k,\\
  & \Rk(X) = 1, \\
  & X \succeq 0, \\
  & z \in \{0,1\}^n.
 \end{aligned}
\end{equation}
We then relax the non-convex rank constraint to arrive at the following
MISDP formulation:
\begin{equation}\label{MISDP}\tag{MISDP}
 \begin{aligned}
  \min \quad & \Tr(A\T A X) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \\
  & \sum_{i=1}^n z_i \leq k, \\
  & X \succeq 0, \\
  & z \in \{0,1\}^n.
 \end{aligned}
\end{equation}
In the following, we will show that \eqref{MISDP} actually produces the
exact value of $\alpha_k$. We will use the following result about the
relationship between the rank of solutions of semidefinite programs and the
dimension of corresponding faces of the feasible set~$S$, i.e., subsets $F
\subseteq S$ such that $\tfrac{1}{2} (p + q) \in F$ for $p, q \in S$ implies
$p,q \in F$:

\begin{theorem}[Pataki~\cite{pat98}]\label{patakiLemma}
  Let $X \in F$, where $F$ is a face of
  \begin{equation*}
    S \define \{X \in S_n \suchthat X \succeq 0,\; \Tr(A_i X) = b_i\; \text{ for } i = 1, \dots, \tilde{m}\}
  \end{equation*}
  for symmetric matrices $A_i \in S_n$, $i = 1, \dots, \tilde{m}$. Then
  \begin{equation*}
    \tfrac{1}{2}\Rk(X)\cdot(\Rk(X)+1) \leq \tilde{m} + dim(F).
  \end{equation*}
\end{theorem}
